---
title: "Warp Breaks"
author: "Mark Pratley"
date: "26 January 2016"
autosize: true
output: 
  slidy_presentation: 
    self_contained: no
--- 
 
```{r, echo=F, warning=F, message=F}
library(ggplot2)
library(dplyr)
library(knitr)
source("../ggplot_themes.R")

wb <- warpbreaks

# tidy up for better graphing
wb <- wb %>% 
        rename(Breaks=breaks, Wool = wool, Tension=tension) %>% 
        mutate(Tension = factor(Tension, labels = c("Low", "Medium", "High")))

```

## Warp Breaks

This is a practice presentation using the [warpbreaks](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/warpbreaks.html) dataset.

The warpbreaks dataset gives the number of breaks in Yarn during Weaving.

It uses fixed lengths of yarn with two different types of wool and varying tensions.

This presentation will investigate which type of wool is recommended for different tensions.


## The Warpbreaks Dataset

The Warpbreaks  dataset has 54 observations with 3 variables:

* Breaks
    + numeric -	The number of breaks

* Wool
    + factor - The type of wool (A or B)
    
* Tension
    + factor - The level of tension (Low, Medium, High)
    
## Examining All Breaks

```{r, echo=F}
# For ggplot use a bin width of 8 

# wb %>% 
#   ggvis(x=~breaks) %>% 
#   layer_histograms(width = input_slider(1, 35, step = 1, label = "Bin Width"), 
#                    center = 10, 
#                    fill := "#E74C3C") 

ggplot(wb, aes(x=Breaks)) +
      geom_histogram(binwidth=7, fill="#c0392b", alpha=0.75) +
      fte_theme() + 
      labs(title="All Breaks", x="# of Breaks", y="Count") +
      scale_x_continuous(breaks=seq(0,80, by=5)) +
      geom_hline(yintercept=0, size=0.4, color="black")
```

## Breaks Grouped by Tension

```{r, echo=F}
wb %>% 
  ggplot(aes(x=Breaks, colour=Tension)) +
    geom_density(size=1)

```

## Breaks Grouped by Tension

```{r, echo=F}
wb %>% 
  ggplot(aes(x=Tension, y=Breaks, colour=Tension)) +
    geom_boxplot(size=1.0) + 
    fte_theme() + 
    labs(x="Tension", y="# No Breaks") +
    geom_hline(yintercept=0, size=0.4, color="black")

```


<!-- ## Breaks Grouped by Tension -->

<!-- ### BoxPlot -->
<!-- ```{r, echo=F, warning=F, message=FALSE} -->
<!-- # Boxplot with hover values -->
<!-- # all_values <- function(x) { -->
<!-- #   if(is.null(x)) return(NULL) -->
<!-- #   paste0(names(x), ": ", format(x), collapse = "<br />") -->
<!-- # } -->
<!-- # wb %>%  -->
<!-- #     ggvis(x=~Tension, y=~Breaks, fill=~Tension ) %>%  -->
<!-- #     layer_boxplots() %>%  -->
<!-- #     add_tooltip(all_values, "hover") %>%  -->
<!-- #     hide_legend("fill") -->

<!-- # Replace with ggplot -->
<!-- wb %>%  -->
<!--   ggplot(aes(x=Tension, y=Breaks, fill=Tension)) + geom_boxplot() -->
<!-- ``` -->

## Tension Conclusions

```{r, echo=F, warning=F, message=FALSE}
wb %>% 
  group_by(Tension) %>% 
  summarise(Num=n(), 
            min=min(Breaks), 
            max=max(Breaks), 
            mean=mean(Breaks),
            median=median(Breaks),
            sd=sd(Breaks)
            ) %>% 
  as.data.frame() %>% 
  kable()
```

From the graphs and table statistics:

Higher tension means less breaks.

Low tension gives more breaks and has higher unpredictability.

## Comparing Breaks with Wool Type

```{r, echo=F}
wb %>% 
  ggplot(aes(x=Breaks, colour=Wool)) +
    geom_density(size=1) +
    fte_theme() + 
    labs(x="# No Breaks", y="Density") +
    theme(legend.justification=c(0,0), legend.position=c(0.6,0.6))
    

```

## Comparing Breaks with Wool Type

```{r, echo=F}
wb %>% 
  ggplot(aes(x=Wool, y=Breaks, colour=Wool)) +
    geom_boxplot(size=1) +
    fte_theme() + 
    labs(x="Wool Type", y="# No Breaks")

```

## Wool Conclusions

```{r, echo=F, warning=F, message=FALSE}
wb %>% 
  group_by(Wool) %>% 
  summarise(Num=n(), 
            min=min(Breaks), 
            max=max(Breaks), 
            mean=mean(Breaks),
            median=median(Breaks),
            sd=sd(Breaks)
            ) %>% 
  as.data.frame() %>% 
  kable()
```

From the statisitics table and the graphs, we can see that overall wool type B has less breaks with lower mean and median values.

Whilst wool type A has a larger standard deviation, which along with a lower minimum and higher maximum shows that wool A has much higher variation than type B.

It is interesting to note that the multimodal appearance (multiple peaks) on the density plot may imply that the different wool types have differing sensitivities to tension. We will explore this next.

## Comparing Breaks with Wool Type and Tension

```{r, echo=F}
# Group and colour by Wool Type & Tension
wb %>% 
  group_by(Group=interaction(Wool, Tension)) %>% 
  ggplot(aes(x=Breaks, group=Group, colour=Tension)) +
    geom_density(size=1, aes(linetype=Wool))+ #, color=sex))+
    fte_theme() + 
    labs(x="# No Breaks", y = "Density") +
    theme(legend.justification=c(0,0), legend.position=c(0.5,0.4)) +
    ggtitle("Number of Warp Breaks with Wool Type and Tension")

```

## Comparing Breaks with Wool Type and Tension

```{r, echo=F}
wb %>% 
  ggplot(aes(x=Tension, y=Breaks, colour=Wool)) +
    geom_boxplot(size=1, 
                 position = position_dodge(width = 0.5), 
                 width = 0.5, 
                 alpha = 0.5
                 )

```

## Mean Warp Breaks by Tension and Wool
```{r, echo=F}
wb %>% 
    group_by(Tension, Wool) %>% 
    summarise(Num=n(), 
              min=min(Breaks), 
              max=max(Breaks), 
              mean=mean(Breaks),
              median=median(Breaks),
              sd=sd(Breaks)
    ) %>% 
    as.data.frame() %>% 
    ggplot(aes(x=Tension, y=mean, fill=Wool)) +
        geom_bar(position = "dodge", stat="identity") +
        labs(y = "Mean Warp Breaks")

```

## Final Table of Statistics for Breaks
```{r, echo=F}
wb %>% 
    group_by(Tension, Wool) %>% 
    summarise(Num=n(), 
              min=min(Breaks), 
              max=max(Breaks), 
              mean=mean(Breaks),
              median=median(Breaks),
              sd=sd(Breaks)
    ) %>% 
    as.data.frame() %>% 
    kable()
```

Generally wool B is better, with the best combination being Wool B with high tension.

The worst combination is Wool A with low tension, although Wool A is better if you exclusively use medium tension.

## Conclusions

If you are only using medium tension then you should choose wool A.

Otherwise, always choose wool B.

If you can choose both wool and tension then you should choose Wool B with High Tension.